Letter: Performance analysis of \\ high-resolution ice sheet simulations

### 2022-06-28 paper/perfmod.tex
Add an ISSM ref.  People are really cagey about their time steps.

For an explicit SIA model the well-known stability restriction is $\Delta t < O(D^{-1} \Delta x^2)$ \citep{Bueleretal2005,HindmarshPayne1996} where $D$ is a representative diffusivity value (from equation \eqref{eq:siamasscontinuity}).  For Stokes dynamics the stability of explicit time-stepping is largely unexplored in any precise sense, but we propose that an advective restriction $\Delta t < O(U^{-1} \Delta x)$, for some representative horizontal velocity scale $U$, or computed from modeled ice velocities, represents the \emph{optimistic} paradigm.  The corresponding \emph{pessimistic} paradigm enforces $\Delta t < O(D^{-1} \Delta x^2)$, using (otherwise diagnostic) diffusivity values computed in the SIA manner.

Explicit time-stepping with hybrid and higher-order schemes is better-studied than for Stokes dynamics, especially over horizontal resolutions relevant to whole ice sheets.  Some hybrid schemes apply the pessimistic time step as an adaptive restriction \citep{Winkelmannetal2011}.  The optimistic time step is supported for a certain higher-order DIVA scheme \citep[see equation (52)]{Robinsonetal2022}, but practical Greenland simulations in the same work actually suggest an intermediate power $\Delta t = O(\Delta x^{1.6})$ (Figure 3(a)).

>>>>

For an explicit SIA model the well-known stability restriction is $\Delta t < O(D^{-1} \Delta x^2)$ \citep{Bueleretal2005,HindmarshPayne1996} where $D$ is a representative diffusivity value (from equation \eqref{eq:siamasscontinuity}).  For Stokes or other membrane-stress-resolving dynamics the stability of explicit time-stepping is largely unexplored in any precise sense, but an advective restriction $\Delta t < O(U^{-1} \Delta x)$, for some representative horizontal velocity scale $U$, or computed from modeled ice velocities \citep[for example]{Fischleretal2022}, represents the \emph{optimistic} paradigm.  The corresponding \emph{pessimistic} paradigm enforces $\Delta t < O(D^{-1} \Delta x^2)$, using (otherwise diagnostic) diffusivity values computed in the SIA manner.

Explicit time-stepping with hybrid and higher-order schemes is better-studied than for Stokes dynamics, especially over horizontal resolutions relevant to whole ice sheets.  Some hybrid schemes apply the pessimistic time step as an adaptive restriction \citep{Winkelmannetal2011}, but some models apparently require the user to choose a time step through trial and error in some circumstances \citep[for example]{Fischleretal2022,Robinsonetal2022}.  The optimistic time step is supported for a certain higher-order DIVA scheme \citep[see equation (52)]{Robinsonetal2022}, but practical Greenland simulations in the same work actually suggest an intermediate power $\Delta t = O(\Delta x^{1.6})$ (Figure 3(a)), and thus some other manner of determining the time step.
